Decorrelating Errors in Quantum Gates by Random Gate Synthesis

### 2018-08-01 decorrelation/decorrelation.tex
one more pass through

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=}
%\hypersetup{citecolor={blue}, colorlinks={true}, filecolor={blue},
%   linkcolor={blue}, pdfauthor=, pdfkeywords= pdfsubject=, pdftitle=,
%   urlcolor{blue}}

\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

@@

\author{Anthony Polloreno}
% \email[Email: ]{apolloreno}
\affiliation{Rigetti Computing, Berkeley, CA}

@@

%As an example, many models assume that errors are local, however systems can have spatially correlated error on more than two qubits that arise from multi-qubit terms in the Hamiltonian, such as:
%\begin{equation}\label{eq:0}
%  H(t) = \sum_{n}\sum_{i}\sum_{j=x,y,z} a^{\textbf{i}}_{\textbf{j}}(t)(\sigma_{j_1}^{i_1}\otimes\sigma_{j_2}^{i_2}\otimes ... \otimes \sigma_{j_n}^{i_n})
%\end{equation}
%where $a^{\textbf{i}}_{\textbf{j}}$ is a stochastic process. In a standard nearest-neighbors topology, like those found in super-conducting qubit architectures, these terms might arise from an effective model where high-frequency degrees of freedom are intergrated out to yield high-weight operators. These noise channels are one way in which the assumptions used in proofs of thresholds may be violated.

There are many ways in which devices can behave in non-Markovian ways,\cite{Kelly2018, BlumeKohout2017}, and even if a system is Markovian, many models make simplifying assumptions about the noise on that system - such as positing that it has a simple structure. For instance Pauli channels are often chosen\needcite due to their classical simulability\needcite. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown, potentially channel depdenent error, coherent errors can be averaged away into incoherent error, at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing. Some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise.

%A common feature of these routines is that take advantage of the fact that processes are generally enacted as a sequence of gates. By inserting a potentially expensive classical pre-compilation step, they can exploit sequence structure of a process to average away coherent error. In the case of \cite{Campbell2017}, the analysis takes the point of view of treating Pauli and Clifford gates as being free, which is out of reach of modern quantum computers\needcite. In \cite{Wallman2016} and the implementation in \cite{Ware2018}, the random frames are implemented in each sequence by precompiling the inserted Pauli frame rotations with the Clifford operations. In general, this cannot be done efficiently for non-Clifford operations, and there for this routine is demanding given the severe circuit depth constraints of modern quantum computers.\needcite

In this Letter, we take a different approach to address these problems at the gate synthesis level. We propose to inject additional decorrelating randomness into the system through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that all approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

Where $DPN[\alpha](\rho)$ is a \textit{generalized} depolarizing noise channel with strength $\alpha$. (For the rest of this paper, we will refer to them just as depolarizing channels.) Such a channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one. This means that on average, the unitary approximations implement the target unitary followed by a small depolarizing channel. The task of constructing the BCSs will fall to optimal control.

This techique can be used to turn coherent error into incoherent error therefore correlated error into uncorrelated error. One benefit of this is that non-Markovian noise will be made to be more Markovian. This is particularly useful because many routines exist that can assess the quality of a gateset, however in the presence of non-Markovianity most of them become unreliable.

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs change the error to a depolorizing channel, the correlations in the noise can be reduced, and the process can be made much closer to Markovian, and as we will show, this comes at no real cost to the fidelity of the implemented gate. Finally, unlike in \cite{Wallman2016, Campbell2017, Ware2018} we note that this method of coherent error mitigation is cheap to implement - it only requires a random choice of pulse definition at run time, in addition to the extra storage required to store the pulse shapes - no classical preprocessing or runtime frame tracking is required.

@@

\begin{figure}
  \centering
  \includegraphics[width=.75\columnwidth]{simple_example.png}
  \caption{An example of a BCS. The red path is a $\pi$ pulse over-rotating clockwise, while the blue path is a $\pi$ pulse over-rotating counter-clockwise. The purple path is the average of the two.}
  \label{fig:simple_example}
\end{figure}
\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BCSs can be done in a variety of ways, using any quantum optimal control technique \cite{Caneva2011, Machnes2018} to find a family of controls. For simplicity in this paper we use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $u_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}u_{ij}H_{ij}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U_n$, and our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U_n\}$.

To optimize this cost function we can perform the following standard update loop for some threshold value $\varepsilon > 0$ and step size $\delta > 0$:
\begin{algorithm}[H]
\floatname{algorithm}
  \caption{\textsc{\textbf{Gradient Ascent}}}
  \begin{algorithmic}
    \While{$J(U_n) < (1-\varepsilon$)}
    \State $u_{ij} \rightarrow u_{ij} + \delta\frac{\partial J(U)}{\partial u_{ij}}$
    \For{$1 \leq i \leq n$}
    \State $U_i \rightarrow \exp\{-i\Delta t(H_0 + \sum_{i=0}^{n}u_{ij}H_i)\}$
    \EndFor
    \State $U \rightarrow \prod_1^nU_i$
    \EndWhile 
  \end{algorithmic}
\end{algorithm}

@@

Given such a Hamiltonian, a threshold approximation fidelity $\mathcal{F}$, and a target gate $U_T$, we can use the technique described in \ref{GRAPE} to generate candidate pulse sequences that enact unitaries $U_i$ that approximate $U_T$, such that $Tr(U_T^{\dagger}U_i) \ge \mathcal{F}$.

In this paper we have modified the update step in \ref{eq:update} to instead use approximately the following gradient:
\begin{align}\label{quadrature}
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed. This technique has been used in previous works such as \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors, and we approximate this integral in this paper by using Gaussian quadrature.\needcite Doing this ensures that the family of controls produced by our routine, $U_i$, perform moderately well over a range of control errors that might occur.

Concretely we consider a Hamiltonian of the following form:
\begin{equation}\label{eq:2}
  H(t) = \delta_0H_0 + \sum_{i=1}^n (1 + \delta_i)c_i(t)H_i
\end{equation}
for control Hamiltonians $H_i$, free evolution Hamiltonian $H_0$ and random variables $\delta_i$, that model some small uncertainty in parameters in the Hamiltonian. Such a model might describe a superconducting qubit quantum processor where control amplitudes for the RF pulses vary over time, or a trapped ion quantum computer where the intensity, frequency, and phase of the laser might drift over time.\cite{Lekitsch2017} Correlations between different $\delta_i$ might arise if two of the controls have the same noise source. Examples of shared noise sources include $X$ and $Y$ gates in superconducting qubit architectures might use the same AWG and pulse envelope, and diurnal temperature drift of control electronics. \needcite

\subsection{BCS Approximation}
After using GRAPE or another optimal control routine to synthesize a collection of controls, we must find the weights $w_i$ such that the collection of controls form a BCS as described in (\ref{eq:1}). To do this, for each control $U_i$ we find the unitary error channel $\mathcal{E}_i$ such that $\mathcal{E}_iU_i=U_T$, where $U_T$ is the target gate. If we consider the Pauli-Liouville representation\cite{Kimmel2014} of this error channel, the diagonal terms are the \textit{stochastic} terms that arise from classical uncertainty, while the off-diagonal terms may more generally arise from \textit{coherent} operations. In particular, we see that we can write a convex sum over these channels as:
\begin{align}
 \frac{1}{N} \sum^N_{i=1} w_i \mathcal{E}^{\dagger} (U_T\rho U_T^{\dagger}) \mathcal{E}
\end{align}
Now, to approximate a depolarizing channel we define our optimal control problem to be the following, which minimizes the off-diagonal terms:
\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{w_0, ..., w_N}{\textbf{minimize}} \{\sum_{i\neq j}^N|\sigma_i\Lambda(\sigma_j)|^2\}\\
    &\textbf{where}\ \Lambda(\sigma_j) := \sum^N_{i=1}w_i\mathcal{E}_i^{\dagger}\sigma_j\mathcal{E}_i\\
    &\textbf{subject to} \sum_{i=1}^Nw_i = 1
  \end{split}
\end{equation}

@@

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires optimizing over the diamond norm, and in particular does not have the restriction that the optimal channel be decomposable into a given family of controls. Our routine, on the other hand, optimizes over an easy to compute sum. In the next section we give a simple example, followed by numerical results for one qubit and two qubits gates in sections \ref{1Q Gates} and \ref{2Q Gates}.

>>>>

\usepackage{amsfonts,amsmath,amssymb,amsthm}
\usepackage{array,bm,color}
\usepackage{epsfig,graphicx,nomencl,revsymb4-1,upgreek,url}
\usepackage{hyperref}
\usepackage{algorithm}
\usepackage{algpseudocode}
\usepackage{subcaption}
\hypersetup{colorlinks=true, pdfauthor=Kevin C. Young, pdftitle=}
\newcommand{\tr}{{\rm Tr\thinspace}}
\newcommand{\bra}[1]{\ensuremath{\left\langle{#1}\right\vert}}
\newcommand{\ket}[1]{\ensuremath{\left\vert{#1}\right\rangle}}
\newcommand{\braket}[2]{\left\langle #1 | #2 \right\rangle}
\newcommand{\ketbra}[2]{\left| #1 \right\rangle\!\!\!\,\left\langle #2 \right|}
\newcommand{\abs}[1]{\left\vert #1 \right\vert}
\newcommand{\expect}[1]{\ensuremath{\left\langle{#1}\right\rangle}}
\newcommand{\timeorder}{\ensuremath{\underset{\leftarrow}{\mathcal{T}}}}
\newcommand{\ident}{{\mathbb1}}
\newcommand{\order}[1]{\mathcal{O}\left( #1 \right)}
\newcommand{\diag}[1]{\mathrm{diag}\{#1\}}
\newcommand{\trans}[1]{#1^\mathsf{T}}
\newcommand{\T}{\mathsf{T}}

@@

\author{Anthony Polloreno}
\email[Email: ]{anthony@rigetti.com}
\affiliation{Rigetti Computing, Berkeley, CA}

@@

There are many ways in which device noise can be non-Markovian ways,\cite{Kelly2018, BlumeKohout2017}, even when the noise in a system is Markovian, many models make simplifying assumptions about the noise on that system. For instance Pauli channels are often used to model systems\needcite due to their classical simulability\needcite, despite their unrealistic form. Some authors have then taken the approach of using these easily-understood models to upper bound the error in a real system by \textit{honestly} representing them \cite{Puzzuoli2014}. This approach is correct and rigorous, but scales poorly.

Other authors have tried to address these problems \cite{Wallman2016, Campbell2017, Heim2016} by using circuit composition to remove correlated noise. Given a process that is generated by a sequence of channels with some unknown, potentially channel dependent error, coherent errors can be averaged away into incoherent error, at the cost of additional compilation steps, which may in general increase circuit depth and require more sophisticated classical preprocessing. Some of these routines have been demonstrated experimentally \cite{Ware2018}, and been shown to reduce the coherence of noise.

In this Letter, we take a different approach to address these problems at the gate synthesis level. We propose to inject additional decorrelating randomness into the system through the use of \emph{balanced control solutions} (BCSs). BCSs are families of control solutions that approximate the same target gate, but with balanced errors for any given instance of the noise Hamiltonian. That is, for a target gate, $U_T$, we seek a family of control solutions, $c_i(t)$, each implementing an approximation $U_i$ to the target gate, such that the family of unitary approximations is \emph{balanced}. A balanced family is one which satisfies, for some small $\alpha$,
\begin{equation}\label{eq:1}
  \frac{1}{N}\sum_{i=1}^N \omega_i U_i \rho U_i^\dagger = DPN[\alpha]\left(U_T \rho U_T^\dagger \right)
\end{equation}

Where $DPN[\alpha](\rho)$ is a \textit{generalized} depolarizing noise channel with strength $\alpha$. (For the rest of this paper, we will refer to them as just depolarizing channels.) Such a channel is defined as:
\begin{equation}\label{eq:2}
  DPN[\alpha](\rho) \rightarrow (1-\alpha)\rho + \alpha\sum p_i \sigma_i\rho\sigma_i
\end{equation}
with $p_i$ summing to one. This means that on average, the unitary approximations implement the target unitary followed by a small depolarizing channel. The task of constructing the BCSs will fall to optimal control.

This techique can be used to turn coherent error into incoherent errorr, correlated error into uncorrelated error, and therefore non-Markovian noise into more Markovian noise. This is particularly, since many routines exist that can assess the quality of a gateset, however in the presence of non-Markovianity most of them become unreliable.

For example, randomized benchmarking and tomography will report incorrect answers without any syndrome \cite{Merkel2013}, and gateset tomography will report that the gateset failed to be Markovian, but will fail to diagnose in what way it was non-Markovian. Because BCSs can make error channels look more depolarizing, the correlations in the noise can be reduced, and the process can be made much closer to Markovian. In addition, we will show that this comes at no real cost to the fidelity of the implemented gate. Finally, unlike in \cite{Wallman2016, Campbell2017, Ware2018} this method of coherent error mitigation is cheap to implement - it only requires a random choice of pulse definition at run time, in addition to the extra storage required to store the pulse shapes - no classical preprocessing or runtime frame tracking is required.

@@

\begin{figure}
  \centering
  \includegraphics[width=.75\columnwidth]{simple_example.png}
  \caption{THIS FIGURE SHOULD BE REPLACED}
  \label{fig:simple_example}
\end{figure}
\section{Optimal Control Problems}\label{ocp}
\subsection{Random Gate Synthesis}
Generating BCSs can be done in a variety of ways, using any quantum optimal control technique \cite{Caneva2011, Machnes2018} to find a family of controls. For simplicity in this paper we use the GRAPE algorithm to generate candidate pulseshapes to approximate the target gate. First described in \cite{Khaneja2005}, the GRAPE (GRadient Ascent Pulse Engineering) algorithm is a technique for finding piecewise constant control sequences that approximate a desired unitary, $U_T$. Defining our uncontrolled Hamiltonian as $H_0$, our control Hamiltonians as $H_{i\neq 0}$, and our \textit{control matrix} $u_{ij}$ as containing control amplitude associated with the $i^{th}$ time step and the $j^{th}$ hamiltonian, we can write our approximate unitary at any timestep as
\begin{equation}\label{eq:3}
  U_i = \exp\{-i\Delta t(H_0 + \sum_{j=1}^{n}u_{ij}H_{j}\}
\end{equation}
Then, to measure the simularity of our approxiate unitary $U_n$ to our target unitary $U_T$, we can define a cost function $J(U) = Tr\{U_T^{\dagger}U_n\}$.

To optimize this cost function we can perform the following standard update loop for some threshold value $\varepsilon > 0$ and step size $\delta > 0$:
\begin{algorithm}[H]
\floatname{algorithm}
  \caption{\textsc{\textbf{Gradient Ascent}}}
  \begin{algorithmic}
    \While{$J(U_n) < (1-\varepsilon$)}
    \State $u_{ij} \rightarrow u_{ij} + \delta\frac{\partial J(U)}{\partial u_{ij}}$
    \For{$1 \leq i \leq n$}
    \State $U_i \rightarrow \exp\{-i\Delta t(H_0 + \sum_{i=0}^{n}u_{ij}H_j)\}$
    \EndFor
    \State $U \rightarrow \prod_{i=1}^nU_i$
    \EndWhile 
  \end{algorithmic}
\end{algorithm}

@@

In our numerical results in Section \ref{1Q Gates} and \ref{2Q Gates}, we have modified the update step in \ref{eq:update} to instead use approximately the following gradient:
\begin{align}\label{quadrature}
\int p(\vec{\delta})\frac{\partial J(U(\vec{\delta}))}{\partial u_{ij}} d\vec{\delta}
\end{align}
with $p(\vec{\delta})$ Gaussian distributed. This technique has been used in previous works such as \cite{Goerz2014} to ensure that the optimal control results are robust over a wide range of errors, and we approximate this integral in this paper by using Gaussian quadrature.\needcite Doing this ensures that the family of controls produced by our routine, $U_i$, perform moderately well over a range of control errors that might occur.

Concretely we consider a Hamiltonian of the following form:
\begin{equation}\label{eq:2}
  H(t) = \delta_0H_0 + \sum_{i=1}^n (1 + \delta_i)c_i(t)H_i
\end{equation}
for control Hamiltonians $H_i$, free evolution Hamiltonian $H_0$ and random variables $\delta_i$, that model some small uncertainty in parameters in the Hamiltonian. Such a model might describe a superconducting qubit quantum processor where control amplitudes for the RF pulses vary over time, or a trapped ion quantum computer where the intensity, frequency, and phase of the laser might drift over time.\cite{Lekitsch2017} Correlations between different $\delta_i$ might arise if two of the controls have the same noise source. Examples of shared noise sources include $RX$ and $RY$ gates in superconducting qubit architectures might use the same AWG and pulse envelope, and diurnal temperature drift of control electronics. \needcite

\subsection{BCS Approximation}
After using GRAPE or another optimal control routine to synthesize a collection of controls, we must find the weights $w_i$ such that the collection of controls form a BCS as described in (\ref{eq:1}). To do this, for each control $U_i$ we find the unitary error channel $\mathcal{E}_i$ such that $\mathcal{E}_iU_i=U_T$, where $U_T$ is the target gate. If we consider the Pauli-Liouville representation\cite{Kimmel2014} of this error channel, the diagonal terms are the \textit{stochastic} terms that arise from classical uncertainty, while the off-diagonal terms may more generally arise from \textit{coherent} operations. In particular, we see that we can write a convex sum over these channels as:
\begin{align}
 \frac{1}{N} \sum^N_{i=1} w_i \mathcal{E}^{\dagger} (U_T\rho U_T^{\dagger}) \mathcal{E}
\end{align}
Now, to approximate a depolarizing channel we define our optimal control problem to be the following, which minimizes the off-diagonal terms:
\begin{equation}\label{eq:minimization}
  \begin{split}
    &\underset{w_0, ..., w_N}{\textbf{minimize}} \{\sum_{\substack{i,j \\ i\neq j}}^N|\sigma_i\Lambda(\sigma_j)|^2\}\\
    &\textbf{where}\ \Lambda(\sigma_j) := \sum^N_{i=1}w_i\mathcal{E}_i^{\dagger}\sigma_j\mathcal{E}_i\\
    &\textbf{subject to} \sum_{i=1}^Nw_i = 1
  \end{split}
\end{equation}

@@

Previous authors have considered minimizing the diamond distance to the nearest Pauli or Clifford Channel \cite{Magesan2013}, and while this gives a good theoretical framework, it requires optimizing over the diamond norm, and in particular does not have the restriction that the optimal channel be decomposed into a given family of controls. Our routine, on the other hand, optimizes over an easy to compute sum, and produces a channel defined in terms of our collection of pulseshapes.
